---
title: "DataAnalysis"
author: "Xavier"
date: "2023-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#6. Exploratory Analysis without Covariates
##6.1. Simple Tables
```{r}
table(df_no_na$laid_off, df_no_na$dum_35_inp)
table(df_no_na$laid_off, df_no_na$dum_40_inp)
table(df_no_na$laid_off, df_no_na$dum_45_inp)

library(ggplot2)

library(ggplot2)

# Creating the continuous plot with a logistic regression line
ggplot(data = df_no_na, aes(x = rosenberg_inp, y = laid_off)) +
  geom_jitter(width = 0.05, height = 0)+
  geom_point() +  # Add points
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  theme_minimal() +  # Optional: adds a minimalistic theme
  labs(x = "Rosenberg Score (Imputed)", y = "Laid Off (Corrected)",
       title = "Relationship between Rosenberg Score and Layoff Status",
       subtitle = "Logistic Regression Line")

```

##6.2. Kaplan Meier Curves
Fitting a Kaplan Meier Curve
```{r}
library(survival)
library(survminer)

surv_object <- Surv(time = df_no_na$time, event = df_no_na$laid_off)

km_fit <- survfit(surv_object ~ 1)  # '~ 1' indicates no stratification

ggsurvplot(km_fit, data = df_no_na, conf.int = TRUE, 
           title = "Kaplan-Meier Survival Curve",
           xlab = "Time", ylab = "Survival Probability")

remove(surv_object, km_fit)


```

Splitting the curve by sub-35 dummy: 
```{r}
surv_object <- Surv(time = df_no_na$time, event = df_no_na$laid_off)
km_fit_stratified <- survfit(surv_object ~ dum_35_inp, data = df_no_na)
ggsurvplot(km_fit_stratified, data = df_no_na, conf.int = TRUE,
           palette = c("blue", "red"),   # Optional: specify colors
           title = "Stratified Kaplan-Meier Survival Curves",
           xlab = "Time", ylab = "Survival Probability",
           pval = TRUE)  # Optional: add a p-value for the comparison

remove(surv_object, km_fit_stratified)
```
Sub-40
```{r}
surv_object <- Surv(time = df_no_na$time, event = df_no_na$laid_off)
km_fit_stratified <- survfit(surv_object ~ dum_40_inp, data = df_no_na)
ggsurvplot(km_fit_stratified, data = df_no_na, conf.int = TRUE,
           palette = c("blue", "red"),   # Optional: specify colors
           title = "Stratified Kaplan-Meier Survival Curves",
           xlab = "Time", ylab = "Survival Probability",
           pval = TRUE)  # Optional: add a p-value for the comparison

remove(surv_object, km_fit_stratified)
```

Sub-45
```{r}
surv_object <- Surv(time = df_no_na$time, event = df_no_na$laid_off)
km_fit_stratified <- survfit(surv_object ~ dum_45_inp, data = df_no_na)
ggsurvplot(km_fit_stratified, data = df_no_na, conf.int = TRUE,
           palette = c("blue", "red"),   # Optional: specify colors
           title = "Stratified Kaplan-Meier Survival Curves",
           xlab = "Time", ylab = "Survival Probability",
           pval = TRUE)  # Optional: add a p-value for the comparison

remove(surv_object, km_fit_stratified)
```

##6.3. Piecewise Constant Model without Covariates
Intervals are years

using different dummies:

```{r}
library(survival)

# Fit the models
# creating yearly intervals
df_no_na$time_interval <- cut(df_no_na$time, breaks = seq(0, max(df_no_na$time), by = 1), include.lowest = TRUE, right = FALSE)


m1_sub_35 <- coxph(Surv(time, laid_off) ~ dum_35_inp + strata(year), data = df_no_na, robust = TRUE)

summary(m1_sub_35) #interpretation: if m1_sub_35 is equal to 1, this corresponds to a  higher risk in getting fired compared to a higher self-esteem. This effect is statistically significantly different from 0. 


m1_sub_40 <- coxph(Surv(time, laid_off) ~ dum_40_inp + strata(year), data = df_no_na, robust = TRUE)

summary(m1_sub_40)

m1_sub_45 <- coxph(Surv(time, laid_off) ~ dum_45_inp + strata(year), data = df_no_na, robust = TRUE)

summary(m1_sub_45)

m1_cont <- coxph(Surv(time, laid_off) ~ rosenberg_inp + strata(year), data = df_no_na)

summary(m1_cont, robust = TRUE)

remove(m1_sub_35, m1_sub_40, m1_sub_45, m1_cont)

```

#7. Gender Effects, No Further Covariates
##7.1. Using Simple Tables 
```{r}
table(df_no_na$female, df_no_na$dum_35_inp)
table(df_no_na$female, df_no_na$dum_40_inp)
table(df_no_na$female, df_no_na$dum_45_inp)

table(df_no_na$female, df_no_na$laid_off)
```

##7.2. Kaplan-Meier Curves 

Using 35 dummy
```{r}
library(ggplot2)

par(mfrow = c(2, 1))

surv_object <- Surv(time = df_no_na_male$time, event = df_no_na_male$laid_off)
km_fit_stratified <- survfit(surv_object ~ dum_35_inp, data = df_no_na_male)

ggsurvplot(km_fit_stratified, data = df_no_na_male, conf.int = TRUE,
           palette = c("blue", "red"),   # Optional: specify colors
           title = "Male",
           xlab = "Time", ylab = "Survival Probability",
           pval = TRUE)  # Optional: add a p-value for the comparison

remove(surv_object, km_fit_stratified)

surv_object <- Surv(time = df_no_na_fem$time, event = df_no_na_fem$laid_off)
km_fit_stratified <- survfit(surv_object ~ dum_35_inp, data = df_no_na_fem)

ggsurvplot(km_fit_stratified, data = df_no_na_fem, conf.int = TRUE,
           palette = c("blue", "red"),   # Optional: specify colors
           title = "Female",
           xlab = "Time", ylab = "Survival Probability",
           pval = TRUE)  # Optional: add a p-value for the comparison

remove(surv_object, km_fit_stratified)
```

using 40 dummy
```{r}
par(mfrow = c(2, 1))

surv_object <- Surv(time = df_no_na_male$time, event = df_no_na_male$laid_off)
km_fit_stratified <- survfit(surv_object ~ dum_40_inp, data = df_no_na_male)

ggsurvplot(km_fit_stratified, data = df_no_na_male, conf.int = TRUE,
           palette = c("blue", "red"),   # Optional: specify colors
           title = "Male",
           xlab = "Time", ylab = "Survival Probability",
           pval = TRUE)  # Optional: add a p-value for the comparison

remove(surv_object, km_fit_stratified)

surv_object <- Surv(time = df_no_na_fem$time, event = df_no_na_fem$laid_off)
km_fit_stratified <- survfit(surv_object ~ dum_40_inp, data = df_no_na_fem)

ggsurvplot(km_fit_stratified, data = df_no_na_fem, conf.int = TRUE,
           palette = c("blue", "red"),   # Optional: specify colors
           title = "Female",
           xlab = "Time", ylab = "Survival Probability",
           pval = TRUE)  # Optional: add a p-value for the comparison

remove(surv_object, km_fit_stratified)
```

using 45 dummy
```{r}
par(mfrow = c(2, 1))

surv_object <- Surv(time = df_no_na_male$time, event = df_no_na_male$laid_off)
km_fit_stratified <- survfit(surv_object ~ dum_45_inp, data = df_no_na_male)

ggsurvplot(km_fit_stratified, data = df_no_na_male, conf.int = TRUE,
           palette = c("blue", "red"),   # Optional: specify colors
           title = "Male",
           xlab = "Time", ylab = "Survival Probability",
           pval = TRUE)  # Optional: add a p-value for the comparison

remove(surv_object, km_fit_stratified)

surv_object <- Surv(time = df_no_na_fem$time, event = df_no_na_fem$laid_off)
km_fit_stratified <- survfit(surv_object ~ dum_45_inp, data = df_no_na_fem)

ggsurvplot(km_fit_stratified, data = df_no_na_fem, conf.int = TRUE,
           palette = c("blue", "red"),   # Optional: specify colors
           title = "Female",
           xlab = "Time", ylab = "Survival Probability",
           pval = TRUE)  # Optional: add a p-value for the comparison

remove(surv_object, km_fit_stratified)
```


##7.3. Piecewise Regressions with Gender (Model A)
Are the results driven by outliers? 
```{r}
A_sub_35_female <- coxph(Surv(time, laid_off) ~ dum_35_inp + strata(year), data = df_no_na_fem, robust = TRUE)

summary(A_sub_35_female)

A_sub_35_male <- coxph(Surv(time, laid_off) ~ dum_35_inp + strata(year), data = df_no_na_male, robust = TRUE)

summary(A_sub_35_male)

A_sub_40_female <- coxph(Surv(time, laid_off) ~ dum_40_inp + strata(year), data = df_no_na_fem, robust = TRUE)

summary(A_sub_40_female)

A_sub_40_male <- coxph(Surv(time, laid_off) ~ dum_40_inp + strata(year), data = df_no_na_male, robust = TRUE)

summary(A_sub_40_male)

A_sub_45_female <- coxph(Surv(time, laid_off) ~ dum_45_inp + strata(year), data = df_no_na_fem, robust = TRUE)

summary(A_sub_45_female)

A_sub_45_male <- coxph(Surv(time, laid_off) ~ dum_45_inp + strata(year), data = df_no_na_male, robust = TRUE)

summary(A_sub_45_male)

remove(A_sub_40_female, A_sub_40_male, A_sub_35_male, A_sub_35_female, A_sub_45_female, A_sub_45_male)

#conclusion: while hazard rates are different, both in female and male the dummy is statistically significant
```

#8. Adding other covariates to piecewise regression 
##8.1 Testing for correlation between covariates and self-esteem

Female:
```{r}
# Select only numeric and binary (dummy) variables
numeric_and_dummy_data <- df_no_na_fem[sapply(df_no_na_fem, function(x) is.numeric(x) || all(x %in% c(0, 1)))]

# Calculate correlations
correlation_matrix <- cor(numeric_and_dummy_data, use = "complete.obs")

correlation_data <- as.data.frame(as.table(correlation_matrix))

# Filter for absolute correlations greater than 0.6, excluding self-correlations
high_correlations <- subset(correlation_data, abs(Freq) > 0.6 & Var1 != Var2)

# View the result
high_correlations #there are no problematic correlations in the female dataset, only logically aantalki (number of children) and d.children (dummy if somebody has at least one child should not be included in the same model)

remove(high_correlations, correlation_data, correlation_matrix, numeric_and_dummy_data)
```
Male:
```{r}
numeric_and_dummy_data <- df_no_na_male[sapply(df_no_na_fem, function(x) is.numeric(x) || all(x %in% c(0, 1)))]

# Calculate correlations
correlation_matrix <- cor(numeric_and_dummy_data, use = "complete.obs")

correlation_data <- as.data.frame(as.table(correlation_matrix))

# Filter for absolute correlations greater than 0.6, excluding self-correlations
high_correlations <- subset(correlation_data, abs(Freq) > 0.6 & Var1 != Var2)

# View the result
high_correlations #there are no problematic correlations in the female dataset, only logically aantalki (number of children) and d.children (dummy if somebody has at least one child should not be included in the same model)

remove(high_correlations, correlation_data, correlation_matrix, numeric_and_dummy_data)
```

##8.2 Model B, education and age
```{r}
B_sub_35_female <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + strata(year), data = df_no_na_fem, robust = TRUE)

summary(B_sub_35_female)

test.ph <- cox.zph(B_sub_35_female) #testing assumptions using Schoenfeld Residuals
test.ph #valid

B_sub_35_male <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + strata(year), data = df_no_na_male, robust = TRUE) 

summary(B_sub_35_male)

test.ph <- cox.zph(B_sub_35_male) #testing assumptions using Schoenfeld Residuals
test.ph #valid

B_sub_40_female <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + strata(year), data = df_no_na_fem, robust = TRUE)

summary(B_sub_40_female)

B_sub_40_male <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + strata(year), data = df_no_na_male, robust = TRUE)

summary(B_sub_40_male)

B_sub_45_female <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + strata(year), data = df_no_na_fem, robust = TRUE)

summary(B_sub_45_female)

B_sub_45_male <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + strata(year), data = df_no_na_male, robust = TRUE)

summary(B_sub_45_male)

remove(B_sub_40_female, B_sub_40_male, B_sub_35_male, B_sub_35_female, B_sub_45_female, B_sub_45_male)

#Same

```

##8.3. Model C: Adding children and partner 
```{r}
C_sub_35_female <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + strata(year), data = df_no_na_fem, robust = TRUE)

summary(C_sub_35_female)

test.ph <- cox.zph(C_sub_35_female) #testing assumptions using Schoenfeld Residuals
test.ph #Be careful Partner is statistically significant (!!!)

C_sub_35_male <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + strata(year), data = df_no_na_male, robust = TRUE) 

summary(C_sub_35_male) 

test.ph <- cox.zph(C_sub_35_male)  #testing assumptions using Schoenfeld Residuals
test.ph #valid

C_sub_40_female <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + strata(year), data = df_no_na_fem, robust = TRUE)

summary(C_sub_40_female)

C_sub_40_male <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + strata(year), data = df_no_na_male, robust = TRUE)

summary(C_sub_40_male)

C_sub_45_female <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + strata(year), data = df_no_na_fem, robust = TRUE)

summary(C_sub_45_female)

C_sub_45_male <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + strata(year), data = df_no_na_male, robust = TRUE)

summary(C_sub_45_male)

remove(C_sub_40_female, C_sub_40_male, C_sub_35_male, C_sub_35_female, C_sub_45_female, C_sub_45_male, test.ph)

#similar results as to Model A and B 
```

##8.4. Model D: Adding occupations
```{r}
D_sub_35_female <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + job_type + strata(year), data = df_no_na_fem, robust = TRUE)

summary(D_sub_35_female)

test.ph <- cox.zph(D_sub_35_female) #testing assumptions using Schoenfeld Residuals
test.ph #valid

D_sub_35_male <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + job_type + strata(year), data = df_no_na_male, robust = TRUE) 

summary(D_sub_35_male) 

test.ph <- cox.zph(D_sub_35_male)  #testing assumptions using Schoenfeld Residuals
test.ph #valid

D_sub_40_female <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + job_type + strata(year), data = df_no_na_fem, robust = TRUE)

summary(D_sub_40_female)

D_sub_40_male <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + job_type + strata(year), data = df_no_na_male, robust = TRUE)

summary(D_sub_40_male)

D_sub_45_female <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + job_type + strata(year), data = df_no_na_fem, robust = TRUE)

summary(D_sub_45_female)

D_sub_45_male <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + job_type + strata(year), data = df_no_na_male, robust = TRUE)

summary(D_sub_45_male)

remove(D_sub_40_female, D_sub_40_male, D_sub_35_male, D_sub_35_female, D_sub_45_female, D_sub_45_male, test.ph)

#similar results, but job_types is more significant for males. 
```

##8.5. Model E: Adding Sectors
```{r}
E_sub_35_female <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + strata(year), data = df_no_na_fem, robust = TRUE)

summary(E_sub_35_female)

test.ph <- cox.zph(E_sub_35_female) #testing assumptions using Schoenfeld Residuals
test.ph #valid

E_sub_35_male <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + strata(year), data = df_no_na_male, robust = TRUE) 

summary(E_sub_35_male) 

test.ph <- cox.zph(E_sub_35_male)  #testing assumptions using Schoenfeld Residuals
test.ph #valid

E_sub_40_female <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + strata(year), data = df_no_na_fem, robust = TRUE)

summary(E_sub_40_female)

E_sub_40_male <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + strata(year), data = df_no_na_male, robust = TRUE)

summary(E_sub_40_male)

E_sub_45_female <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + strata(year), data = df_no_na_fem, robust = TRUE)

summary(E_sub_45_female)

E_sub_45_male <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + strata(year), data = df_no_na_male, robust = TRUE)

summary(E_sub_45_male)

remove(E_sub_40_female, E_sub_40_male, E_sub_35_male, E_sub_35_female, E_sub_45_female, E_sub_45_male, test.ph) #weird sector and job_type only matter for males? 
```

##8.6. Model F: Adding Health

```{r}
F_sub_35_female <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + strata(year), data = df_no_na_fem, robust = TRUE)

summary(F_sub_35_female)

test.ph <- cox.zph(F_sub_35_female) #testing assumptions using Schoenfeld Residuals
test.ph #valid

F_sub_35_male <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + strata(year), data = df_no_na_male, robust = TRUE) 

summary(F_sub_35_male) 

test.ph <- cox.zph(F_sub_35_male)  #testing assumptions using Schoenfeld Residuals
test.ph #valid

F_sub_40_female <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + strata(year), data = df_no_na_fem, robust = TRUE)

summary(F_sub_40_female)

F_sub_40_male <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + strata(year), data = df_no_na_male, robust = TRUE)

summary(F_sub_40_male)

F_sub_45_female <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + strata(year), data = df_no_na_fem, robust = TRUE)

summary(F_sub_45_female)

F_sub_45_male <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + strata(year), data = df_no_na_male, robust = TRUE)

summary(F_sub_45_male)

remove(F_sub_40_female, F_sub_40_male, F_sub_35_male, F_sub_35_female, F_sub_45_female, F_sub_45_male, test.ph) 
```



##8.Y. Adding extra dummy for job_satisfaction 
```{r}
Y_sub_35_female <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + strata(year), data = df_no_na_fem, robust = TRUE)

summary(Y_sub_35_female)

test.ph <- cox.zph(Y_sub_35_female) #testing assumptions using Schoenfeld Residuals
test.ph #valid

Y_sub_35_male <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + strata(year), data = df_no_na_male, robust = TRUE) 

summary(Y_sub_35_male) 

test.ph <- cox.zph(Y_sub_35_male)  #testing assumptions using Schoenfeld Residuals
test.ph #valid

Y_sub_40_female <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + strata(year), data = df_no_na_fem, robust = TRUE)

summary(Y_sub_40_female)

Y_sub_40_male <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + strata(year), data = df_no_na_male, robust = TRUE)

summary(Y_sub_40_male)

Y_sub_45_female <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + strata(year), data = df_no_na_fem, robust = TRUE)

summary(Y_sub_45_female)

Y_sub_45_male <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + strata(year), data = df_no_na_male, robust = TRUE)

summary(Y_sub_45_male)

remove(Y_sub_40_female, Y_sub_40_male, Y_sub_35_male, Y_sub_35_female, Y_sub_45_female, Y_sub_45_male, test.ph) 
```


##8.Z. Adding Extra dummy for job-hindrance 
```{r}
Z_sub_35_female <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + job_hindrance + strata(year), data = df_no_na_fem, robust = TRUE)

summary(Z_sub_35_female)

test.ph <- cox.zph(Z_sub_35_female) #testing assumptions using Schoenfeld Residuals
test.ph #valid

Z_sub_35_male <- coxph(Surv(time, laid_off) ~ dum_35_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + job_hindrance + strata(year), data = df_no_na_male, robust = TRUE) 

summary(Z_sub_35_male) 

test.ph <- cox.zph(Z_sub_35_male)  #testing assumptions using Schoenfeld Residuals
test.ph #valid

Z_sub_40_female <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + job_hindrance + strata(year), data = df_no_na_fem, robust = TRUE)

summary(Z_sub_40_female)

Z_sub_40_male <- coxph(Surv(time, laid_off) ~ dum_40_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + job_hindrance + strata(year), data = df_no_na_male, robust = TRUE)

summary(Z_sub_40_male)

Z_sub_45_female <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + job_hindrance + strata(year), data = df_no_na_fem, robust = TRUE)

summary(Z_sub_45_female)

Z_sub_45_male <- coxph(Surv(time, laid_off) ~ dum_45_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + job_hindrance + strata(year), data = df_no_na_male, robust = TRUE)

summary(Z_sub_45_male)

remove(Z_sub_40_female, Z_sub_40_male, Z_sub_35_male, Z_sub_35_female, Z_sub_45_female, Z_sub_45_male, test.ph) 
```

#9. Using Continuous Rosenberg Scale: Lineair and Non-Lineair 
##9.1. Linear modelds
```{r}
A_cont_lin_female <- coxph(Surv(time, laid_off) ~ rosenberg_inp + strata(year), data = df_no_na_fem, robust = TRUE)

summary(A_cont_lin_female)

A_cont_lin_male <- coxph(Surv(time, laid_off) ~ rosenberg_inp + strata(year), data = df_no_na_male, robust = TRUE)

summary(A_cont_lin_male)

B_cont_lin_female <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + strata(year), data = df_no_na_fem, robust = TRUE)

summary(B_cont_lin_female)

test.ph <- cox.zph(B_cont_lin_female) #testing assumptions using Schoenfeld Residuals
test.ph #valid

B_cont_lin_male <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + strata(year), data = df_no_na_male, robust = TRUE) 

summary(B_cont_lin_male)

C_cont_lin_female <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + strata(year), data = df_no_na_fem, robust = TRUE)

summary(C_cont_lin_female)

test.ph <- cox.zph(C_cont_lin_female) #testing assumptions using Schoenfeld Residuals
test.ph #Be careful Partner is statistically significant (!!!)

C_cont_lin_male <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + strata(year), data = df_no_na_male, robust = TRUE) 

summary(C_cont_lin_male) 

test.ph <- cox.zph(C_cont_lin_male)  #testing assumptions using Schoenfeld Residuals
test.ph #valid


D_cont_lin_female <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + job_type + strata(year), data = df_no_na_fem, robust = TRUE)

summary(D_cont_lin_female)

test.ph <- cox.zph(D_cont_lin_female) #testing assumptions using Schoenfeld Residuals
test.ph #valid

D_cont_lin_male <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + job_type + strata(year), data = df_no_na_male, robust = TRUE) 

summary(D_cont_lin_male) 

test.ph <- cox.zph(D_cont_lin_male)  #testing assumptions using Schoenfeld Residuals
test.ph #valid

E_cont_lin_female <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + strata(year), data = df_no_na_fem, robust = TRUE)

summary(E_cont_lin_female)

test.ph <- cox.zph(E_cont_lin_female) #testing assumptions using Schoenfeld Residuals
test.ph #valid

E_cont_lin_male <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + strata(year), data = df_no_na_male, robust = TRUE) 

summary(E_cont_lin_male) 

test.ph <- cox.zph(E_cont_lin_male)  #testing assumptions using Schoenfeld Residuals
test.ph #valid

F_cont_lin_female <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + strata(year), data = df_no_na_fem, robust = TRUE)

summary(F_cont_lin_female)

test.ph <- cox.zph(F_cont_lin_female) #testing assumptions using Schoenfeld Residuals
test.ph #valid

F_cont_lin_male <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + strata(year), data = df_no_na_male, robust = TRUE) 

summary(F_cont_lin_male) 

test.ph <- cox.zph(F_cont_lin_male)  #testing assumptions using Schoenfeld Residuals
test.ph #valid

Y_cont_lin_female <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + strata(year), data = df_no_na_fem, robust = TRUE)

summary(Y_cont_lin_female)

test.ph <- cox.zph(Y_cont_lin_female) #testing assumptions using Schoenfeld Residuals
test.ph #valid

Y_cont_lin_male <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + strata(year), data = df_no_na_male, robust = TRUE) 

summary(Y_cont_lin_male) 

test.ph <- cox.zph(Y_cont_lin_male)  #testing assumptions using Schoenfeld Residuals
test.ph #valid

Z_cont_lin_female <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + job_hindrance + strata(year), data = df_no_na_fem, robust = TRUE)

summary(Z_cont_lin_female)

test.ph <- cox.zph(Z_cont_lin_female) #testing assumptions using Schoenfeld Residuals
test.ph #invalid job_hindrance correlates 

Z_cont_lin_male <- coxph(Surv(time, laid_off) ~ rosenberg_inp + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + job_hindrance + strata(year), data = df_no_na_male, robust = TRUE) 

summary(Z_cont_lin_male) 

test.ph <- cox.zph(Z_cont_lin_male)  #testing assumptions using Schoenfeld Residuals
test.ph #

remove(test.ph, A_cont_lin_male, B_cont_lin_male, C_cont_lin_male, D_cont_lin_male, E_cont_lin_male, F_cont_lin_male, Y_cont_lin_male, Z_cont_lin_male,
       A_cont_lin_female, B_cont_lin_female, C_cont_lin_female, D_cont_lin_female, E_cont_lin_female, F_cont_lin_female, Y_cont_lin_female, Z_cont_lin_female)
```

##9.2. Checking non-linearity
###Graphically 
```{r}
library(survival)
library(survival)

# Linear model
A_cont_lin_female <- coxph(Surv(time, laid_off) ~ rosenberg_inp + strata(year), data = df_no_na_fem, robust = TRUE)

# Non-linear model (using restricted cubic splines for illustration)
A_cont_nonlin_female <- coxph(Surv(time, laid_off) ~ rcs(rosenberg_inp, 3) + strata(year), data = df_no_na_fem, robust = TRUE)

# Create a new data frame for predictions
rosenberg_range <- data.frame(rosenberg_inp = seq(min(df_no_na_fem$rosenberg_inp, na.rm = TRUE), 
                                                  max(df_no_na_fem$rosenberg_inp, na.rm = TRUE), 
                                                  length.out = 100))

# Predict survival probabilities
surv_pred_lin <- survfit(A_cont_lin_female, newdata = rosenberg_range)
surv_pred_nonlin <- survfit(A_cont_nonlin_female, newdata = rosenberg_range)

# Extract survival probabilities and time points
surv_pred_lin_df <- data.frame(
  rosenberg_inp = rep(rosenberg_range$rosenberg_inp, each = length(surv_pred_lin$time)),
  surv = rep(surv_pred_lin$surv, times = length(rosenberg_range$rosenberg_inp))
)

surv_pred_nonlin_df <- data.frame(
  rosenberg_inp = rep(rosenberg_range$rosenberg_inp, each = length(surv_pred_nonlin$time)),
  surv = rep(surv_pred_nonlin$surv, times = length(rosenberg_range$rosenberg_inp))
)

library(ggplot2)

# Plotting
ggplot() +
  geom_line(data = surv_pred_lin_df, aes(x = rosenberg_inp, y = surv), colour = "blue") +
  geom_line(data = surv_pred_nonlin_df, aes(x = rosenberg_inp, y = surv), colour = "red") +
  labs(x = "Rosenberg Self-Esteem Score", y = "Survival Probability") +
  ggtitle("Predicted Survival Probabilities") +
  theme_minimal()





remove(surv_pred_lin_df, surv_pred_nonlin_df, rosenberg_range, A_cont_lin_female, A_cont_nonlin_female)
```

```{r}
library(survival)
cox_model_factor <- coxph(Surv(time, laid_off) ~ rosenberg_factor + strata(year), data = df_no_na, robust = TRUE)

summary(cox_model_factor)


# Calculate hazard ratios and their 95% CI
hr_and_ci <- exp(coef(cox_model_factor))
conf_int <- exp(confint(cox_model_factor))

# Create a data frame for the hazard ratios and confidence intervals
hr_df <- data.frame(
  Variable = names(hr_and_ci),
  HR = hr_and_ci,
  Lower = conf_int[, "2.5 %"],
  Upper = conf_int[, "97.5 %"]
)

library(ggplot2)

# Plotting hazard ratios with confidence intervals
ggplot(hr_df, aes(x = Variable, y = HR)) +
  geom_point() +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0.2) +
  scale_y_continuous(trans = 'log', name = "Hazard Ratio (log scale)") +
  labs(x = "Variable", title = "Hazard Ratios with Confidence Intervals") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_smooth()
remove(cox_model_factor, hr_df, hr_simplified_df, hr_simplified, hr_and_ci, conf_int)
```

And do the same for the full model: 
```{r}

cox_model_factor <- coxph(Surv(time, laid_off) ~ rosenberg_factor + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + strata(year), data = df_no_na, robust = TRUE)

summary(cox_model_factor)

# Calculate hazard ratios and their 95% CI
hr_and_ci <- exp(coef(cox_model_factor))
conf_int <- exp(confint(cox_model_factor))

# Create a data frame for the hazard ratios and confidence intervals
hr_df <- data.frame(
  Variable = names(hr_and_ci),
  HR = hr_and_ci,
  Lower = conf_int[, "2.5 %"],
  Upper = conf_int[, "97.5 %"]
)

# Filter the data frame to include only variables related to Rosenberg
hr_df <- hr_df[grep("rosenberg_factor", hr_df$Variable), ]


library(ggplot2)

# Plotting hazard ratios with confidence intervals
ggplot(hr_df, aes(x = Variable, y = HR)) +
  geom_point() +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0.2) +
  scale_y_continuous(trans = 'log', name = "Hazard Ratio (log scale)") +
  labs(x = "Variable", title = "Hazard Ratios with Confidence Intervals") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_smooth()
remove(cox_model_factor, hr_df, hr_and_ci, conf_int)

```

And now splitted for male and female: 


```{r}
# Cox model for males
cox_model_factor_male <- coxph(Surv(time, laid_off) ~ rosenberg_factor + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + strata(year), data = df_no_na_male, robust = TRUE)

summary(cox_model_factor_male)
# Hazard ratios and CI for males
hr_and_ci_male <- exp(coef(cox_model_factor_male))
conf_int_male <- exp(confint(cox_model_factor_male))

# Data frame for males
hr_df_male <- data.frame(
  Variable = names(hr_and_ci_male),
  HR = hr_and_ci_male,
  Lower = conf_int_male[, "2.5 %"],
  Upper = conf_int_male[, "97.5 %"],
  Gender = "Male"
)

# Filter for Rosenberg factors
hr_df_male <- hr_df_male[grep("rosenberg_factor", hr_df_male$Variable), ]

# Cox model for females
cox_model_factor_female <- coxph(Surv(time, laid_off) ~ rosenberg_factor + d.high_edu + leeftijd + partner + d.children + job_type + sector + d.mental_health + d.heavy_alcohol + job_satisf + strata(year), data = df_no_na_fem, robust = TRUE)

summary(cox_model_factor_female)

# Hazard ratios and CI for females
hr_and_ci_female <- exp(coef(cox_model_factor_female))
conf_int_female <- exp(confint(cox_model_factor_female))

# Data frame for females
hr_df_female <- data.frame(
  Variable = names(hr_and_ci_female),
  HR = hr_and_ci_female,
  Lower = conf_int_female[, "2.5 %"],
  Upper = conf_int_female[, "97.5 %"],
  Gender = "Female"
)

# Filter for Rosenberg factors
hr_df_female <- hr_df_female[grep("rosenberg_factor", hr_df_female$Variable), ]

# Combine male and female data frames
hr_df_combined <- rbind(hr_df_male, hr_df_female)

# Plotting
ggplot(hr_df_combined, aes(x = Variable, y = HR, color = Gender)) +
  geom_point() +
  geom_errorbar(aes(ymin = Lower, ymax = Upper), width = 0.2) +
  scale_color_manual(values = c("Male" = "blue", "Female" = "red")) +
  scale_y_continuous(trans = 'log', name = "Hazard Ratio (log scale)") +
  labs(x = "Rosenberg Score Group", title = "Hazard Ratios by Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_smooth(se = FALSE, method = "loess")

# Remove temporary variables
rm(cox_model_factor_male, hr_and_ci_male, conf_int_male, hr_df_male,
   cox_model_factor_female, hr_and_ci_female, conf_int_female, hr_df_female,
   hr_df_combined)


```

### Statistically
