---
title: "DataPreperation"
author: "Xavier"
date: "2023-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Rosenberg Scores 
##1.1. Creating a long dataset, with Rosenberg scores when available for each year. 
By running 1.1. + 1.2. + 1.3. + 1.4. you will get self_esteem_final, a long dataset with the values for the rosenberg score per year per participant.  

note: There was no personality survey in 2016

```{r}
# Initialize an empty dataframe to store the combined results
self_esteem <- data.frame()

# Define the suffixes for each year
suffixes <- c("2008" = "a", "2009" = "b", "2010" = "c", "2011" = "d", 
              "2012" = "e", "2013" = "f", "2014" = "g", "2015" = "h", 
              "2017" = "i", "2018" = "j", "2019" = "k", "2020" = "l", 
              "2021" = "m", "2022" = "n")

# Loop through the years 2008 to 2022, excluding 2016
for (year in setdiff(2008:2022, 2016)) {
    # Construct the file name based on the year
    file_name <- paste0("RawData/Personality/Personality_", year, ".dta")
    
    # Read the data
    personality_data <- read_dta(file_name)
    
    # Create a temporary dataframe for the year
    temp_df <- data.frame(
        nomem_encr = personality_data$nomem_encr,
        year = year
    )
    
    # Get the suffix for the year
    suffix <- suffixes[as.character(year)]
    
    # Calculate the Rosenberg score components
    temp_pos <- rowSums(personality_data[,paste0("cp", substr(year, 3, 4), suffix, c("070", "071", "073", "075", "076"))], na.rm = FALSE)
    temp_neg <- rowSums(8 - personality_data[,paste0("cp", substr(year, 3, 4), suffix, c("072", "074", "077", "078", "079"))], na.rm = FALSE)
    
    # Check for NA in either temp_pos or temp_neg and calculate Rosenberg score
    temp_df$rosenberg <- ifelse(is.na(temp_pos) | is.na(temp_neg), NA, temp_pos + temp_neg)
    
    # Combine the temporary dataframe with the main dataframe
    self_esteem <- rbind(self_esteem, temp_df)
}

# View the final combined dataframe
View(self_esteem)

```

##1.2. Filling in NA's
it's clear there are many NA's in 2010, 2012, 2015, 2018. But there are still some people that have made the test in those years. 
```{r}
aggregate(is.na(rosenberg) ~ year, data = self_esteem, FUN = sum)
hist(self_esteem$rosenberg[self_esteem$year==2015], breaks = 49)

```

For 2010, 2012, 2015 and 2018 we will inpute the average of the previous years. If one of those is NA, it will take the other one. For 2012 it will use 2011 and 2009 as corresponding years

```{r}
# Function to calculate the individual-specific average for the two previous years
individual_avg <- function(year, id, data) {
    # Filter for the individual and for the two preceding years
    prev_years <- sort(setdiff((year - 3):(year - 1), c(2016)), decreasing = TRUE)
    individual_data <- data[data$nomem_encr == id & data$year %in% prev_years, ]
    
    # If both years are NA, return NA
    if (all(is.na(individual_data$rosenberg))) {
        return(NA)
    }
    
    # Calculate the mean, excluding NA values
    mean(individual_data$rosenberg, na.rm = TRUE)
}

# Apply the function to fill NA values for specific years
for (year in c(2010, 2012, 2015, 2018)) {
    # Identify rows for the specific year with NA in Rosenberg
    rows_to_update <- which(self_esteem$year == year & is.na(self_esteem$rosenberg))
    
    # Calculate and fill in the average for each individual
    for (row in rows_to_update) {
        id <- self_esteem$nomem_encr[row]
        self_esteem$rosenberg[row] <- individual_avg(year, id, self_esteem)
    }
}

# View the updated dataframe
View(self_esteem)

#check na's 
aggregate(is.na(rosenberg) ~ year, data = self_esteem, FUN = sum)
```

##1.3. Inpute average for missing 2016 survey
For 2016 we input 2015, which was calculated above as the individuals specific average over 2014 and 2013.
We only do this for people are already participating in the survey prior to 2016. 


```{r}
# Identify individuals with observations between 2008 and 2015
individuals <- unique(self_esteem$nomem_encr[self_esteem$year %in% 2008:2015])

# Create a new dataframe for 2016 observations
self_esteem_2016 <- data.frame()

# Loop through each individual and impute the 2015 value for 2016
for (id in individuals) {
    # Find the 2015 Rosenberg score for the individual
    rosenberg_2015 <- self_esteem$rosenberg[self_esteem$nomem_encr == id & self_esteem$year == 2015]
    
    # If there is a 2015 score, create a new observation for 2016
    if (length(rosenberg_2015) > 0 && !is.na(rosenberg_2015)) {
        self_esteem_2016 <- rbind(self_esteem_2016, data.frame(nomem_encr = id, year = 2016, rosenberg = rosenberg_2015))
    }
}

# Combine the new 2016 observations with the original dataframe
self_esteem <- rbind(self_esteem, self_esteem_2016)

# View the updated dataframe
View(self_esteem)

```


##1.4. Quick explanatory analysis and adding dummy 
```{r}
#average Rosenberg per year
aggregate(rosenberg ~ year, data = self_esteem, FUN = mean, na.rm = TRUE)

#average Rosenberg per year
aggregate(rosenberg ~ year, data = self_esteem, FUN = median, na.rm = TRUE)

#check na's 
aggregate(is.na(rosenberg) ~ year, data = self_esteem, FUN = sum)

#mean standard deviation within individual 
individual_variation <- aggregate(rosenberg ~ nomem_encr, data = self_esteem, FUN = sd, na.rm = TRUE)

mean(individual_variation$rosenberg, na.rm = TRUE)

#histogram
hist(self_esteem$rosenberg, breaks = 49)

#dummy, coded 0 as below or equal to 35 (low SE) or 1 as higher than 35
#or 0 if below or equal to 40 etc.
self_esteem$dum_35 <- ifelse(self_esteem$rosenberg <= 35, 0, 1)
self_esteem$dum_40 <- ifelse(self_esteem$rosenberg <= 40, 0, 1)
self_esteem$dum_45 <- ifelse(self_esteem$rosenberg <= 45, 0, 1)


table(self_esteem$dum_35)
proportions(table(self_esteem$dum_35))

table(self_esteem$dum_40)
proportions(table(self_esteem$dum_40))

table(self_esteem$dum_45)
proportions(table(self_esteem$dum_45))

```


##1.5. Cleaning up and Saving Data
```{r} 
remove(temp_df, personality_data, file_name, suffix, suffixes, temp_neg, temp_pos, year, id, individuals, rosenberg_2015, row, rows_to_update, individual_avg, self_esteem_2016, individual_variation)

# Define the file path
file_path <- "ProcessedData/self_esteem_final.csv"

# Check if the ProcessedData folder exists, create it if it doesn't
if (!dir.exists("ProcessedData")) {
  dir.create("ProcessedData")
}

# Save the self_esteem dataframe as a CSV file
write.csv(self_esteem, file_path, row.names = FALSE)

# Print a message to confirm saving
print(paste("File saved as", file_path))

```

